Evaluating the Influence of Education on Current Unemployment Trends

### 2024-04-11 paper/paper.qmd
updated paper

---
title: "Evaluating the Influence of Education on Current Unemployment Trends"
author: 
  - Adrian Ly
thanks: "Code and data are available at: [LINK](https://github.com/AdrianUofT/unemployment-employment-rate)"
date: today
date-format: long
abstract: "This paper investigates the relationship between educational attainment and unemployment rates amidst the evolving dynamics of technological advancement and globalization, utilizing data from the Labour Force Survey spanning 2019 to 2023. Our analysis reveals significant variability in unemployment rates across different educational levels, highlighting the nuanced impact of higher education in an era increasingly dominated by artificial intelligence and outsourcing. These findings underscore the critical need for educational systems to adapt, emphasizing skills that complement technological advancements and align with the global job market. This study illuminates the complex interplay between education, employment, and technological change, offering valuable insights for policymakers and educators in preparing the workforce for the challenges of the future labor market."
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

@@

A significant gap exists in understanding the real-time implications of these global shifts on employment, particularly in how educational attainment intersects with evolving job market demands. This study addresses this gap by utilizing data from the Labor Force Survey, exploring the interplay between education and unemployment, and considering the external factors influencing these relationships.

@@

The paper is structured as follows: Following this introduction, we describe the methodology and data sources used in our analysis in (@sec-data). Subsequent sections present the model we estimate (@sec-model), and results (@sec-results) showcasing the implications of educational attainment on unemployment rates. The paper concludes with a discussion (@sec-discussion) of the findings, limitations, and recommendations for future research and policy considerations.

@@

This paper uses the Statistics of Canada `Unemployment rate, participation rate and employment rate by educational attainment, annual` [@citedataset] dataset. We are using the `R programming language` [@citeR] to conduct our analysis along with `readR` [@citereadR], `lubridate` [@citelubridate], `tidyverse` [@citetidyverse], `dplyr` [@citedplyr], `tidyr` [@citetidyr], `knitr` [@citeknitr], `janitor` [@citejanitor], `scales` [@citescales], `RColorBrewer` [@citeRColorBrewer], `ggplot2` [@citeggplot2], `kableExtra` [@citekableExtra], `here` [@citeHere], `arrow` [@citeArrow], `rstanarm` [@citerstanarm], `modelsummary` [@citemodelsummary], and `lme4` [@citelme4].

@@

The figures and tables from our analysis present a nuanced picture of youth unemployment across different educational strata over a period extending from 2019 to 2023.

@@

From the results of our analysis, two key insights emerge that expand our understanding of the global labor market and its interplay with education.

@@

In evaluating the data and methodology utilized in our analysis, several weaknesses warrant attention, alongside considerations for future data collection efforts aimed at refining our understanding of the labor market dynamics.

>>>>

---
title: "Evaluating the Influence of Education on Current Unemployment Trends"
author: 
  - Adrian Ly
thanks: "Code and data are available at: [LINK](https://github.com/AdrianUofT/unemployment-employment-rate)"
date: today
date-format: long
abstract: "This paper investigates the relationship between educational attainment and unemployment rates amidst the evolving dynamics of technological advancement and globalization, utilizing data from the Labour Force Survey spanning 2019 to 2023. The analysis reveals significant variability in unemployment rates across different educational levels, highlighting the nuanced impact of higher education in an era increasingly dominated by artificial intelligence and outsourcing. These findings underscore the critical need for educational systems to adapt, emphasizing skills that complement technological advancements and align with the global job market. This study illuminates the complex interplay between education, employment, and technological change, offering valuable insights for policymakers and educators in preparing the workforce for the challenges of the future labor market."
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

@@

A significant gap exists in understanding the real-time implications of these global shifts on employment, particularly in how educational attainment intersects with evolving job market demands. This study addresses this gap by utilizing data from the Labor Force Survey (@citedatasetsource), exploring the interplay between education and unemployment, and considering the external factors influencing these relationships.

@@

The paper is structured as follows: Following this introduction, we describe the methodology and data sources used in the analysis in (@sec-data). Subsequent sections present the model we estimate (@sec-model), and results (@sec-results) showcasing the implications of educational attainment on unemployment rates. The paper concludes with a discussion (@sec-discussion) of the findings, limitations, and recommendations for future research and policy considerations.

@@

This paper uses the Statistics of Canada `Unemployment rate, participation rate and employment rate by educational attainment, annual` [@citedataset] dataset. We are using the `R programming language` [@citeR] to conduct the analysis along with `readR` [@citereadR], `lubridate` [@citelubridate], `tidyverse` [@citetidyverse], `dplyr` [@citedplyr], `tidyr` [@citetidyr], `knitr` [@citeknitr], `janitor` [@citejanitor], `scales` [@citescales], `RColorBrewer` [@citeRColorBrewer], `ggplot2` [@citeggplot2], `kableExtra` [@citekableExtra], `here` [@citeHere], `arrow` [@citeArrow], `rstanarm` [@citerstanarm], `modelsummary` [@citemodelsummary], and `lme4` [@citelme4].

@@

The figures and tables from the analysis present a nuanced picture of youth unemployment across different educational strata over a period extending from 2019 to 2023.

@@

From the results of the analysis, two key insights emerge that expand our understanding of the global labor market and its interplay with education.

@@

In evaluating the data and methodology utilized in the analysis, several weaknesses warrant attention, alongside considerations for future data collection efforts aimed at refining our understanding of the labor market dynamics.
